---
title: "AssignmentIV"
author: "Silpa Soni Nallacheruvu (19980824-5287) Hernan Aldana (20000526-4999)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Exercise 4:1

##  Question 1

Report the model selection process briefly. Based on your chosen model, which factors affect the probability of not surviving? Report odds ratios with confidence intervals for the most important variables/factors, and interpret them. Use the variable names from the table (not V3, V4, etc.).

### Approach:

- Model Selection Process:

  - Use a stepwise selection with AIC to identify a parsimonious model.
  
-	Analysis of the Final Model:

  - Extract coefficients, odds ratios, and their 95% confidence intervals for significant variables.
  
  - Ensure variable names are replaced with their descriptions (e.g., Age, Sex, etc.) instead of column names.

- Interpretation:

  - Interpret the results form the AIC, odds ratios, and confidence intervals, to determine the best model.

### Code and results :

```{r echo=FALSE}
library(dplyr)
data_ca4 <- read.csv("data_ca4.csv")

# Rename variables for clarity
colnames(data_ca4) <- c(
  "Patient", "Survival", "Age", "Sex", "Ethnicity", "TreatmentAtAdmission",
  "Cancer", "PreviousKidneyFailure", "Infection", "HeartLungTreatment",
  "BloodPressure", "HeartRate", "AdmittedToICUWithin6Months",
  "TypeOfAdmission", "FractureOfNeckOrSpine", "BloodOxygen",
  "BloodPH", "BloodCarbonDioxide", "BloodBicarbonate",
  "BloodCreatine", "ConsciousnessLevel"
)

# Combine categories for Ethnicity (example)
data_ca4$Ethnicity[data_ca4$Ethnicity > 1] <- 0

# Fit models
m_empty <- glm(Survival ~ 1, family = binomial, data = data_ca4)   # Empty model
m_full <- glm(Survival ~ ., family = binomial, data = data_ca4)    # Full model

# Stepwise model selection using AIC
m_step <- step(m_empty, scope = list(lower = m_empty, upper = m_full), direction = "both", trace = FALSE)

# Summary of the final model
summary(m_step)

```


**Model selection**

- From the output the final logistic regression model includes the following variables: ConsciousnessLevel, TypeOfAdmission, Age, Cancer, Patient, BloodCarbonDioxide, BloodPH, and BloodPressure.

- These variables where selected using a stepwise AIC, which ensures a balance between model complexity and goodness of fit.

**Significant Variables **

- Variables with a p-value < 0.05 are considered significant predictors of survival

  - ConsciousnessLevel
  
  - TypeOfAdmission
  
  - Age
  
  - Cancer
  
  - BloodCarbonDioxide, 
  
  - BloodPH


```{r echo=FALSE}

# Extract Odds Ratios and Confidence Intervals
odds_ratios <- exp(coef(m_step))  # Odds Ratios
conf_intervals <- exp(confint(m_step))  # 95% Confidence Intervals

# Combine results into a data frame for clarity
results <- data.frame(
  Variable = names(odds_ratios),
  OddsRatio = odds_ratios,
  `CI Lower` = conf_intervals[, 1],
  `CI Upper` = conf_intervals[, 2]
)

# Filter significant variables (p-value < 0.05)
significant_vars <- summary(m_step)$coefficients
significant_vars <- significant_vars[significant_vars[, 4] < 0.05, ]

# Final table for reporting
final_results <- results %>%
  filter(Variable %in% rownames(significant_vars))

final_results
```

**Odds Ratios and Confidence Intervals:**


1. ConsciousnessLevel: 

   - Odds Ratio: 13.75 (CI: 4.31-65.28)
   
   - Patients who are unconscious or in a coma have a significantly higher probability of not surviving compared to those who are conscious.
   
2. TypeOfAdmission:

   - Odds Ratio: 21.25 (CI: 4.36-189.15)
   
   - Acute admissions are associated with a significantly higher probability of not surviving compared to non-acute admissions.
   
3. Age:

   - Odds Ratio: 1.04 (CI: 1.01-1.07)
   
   - For each additional year of age, the odds of not surviving increase by 4%.

4. Cancer:

   - Odds Ratio: 10.37 (CI: 1.95-66.54)
   
   - Patients with cancer have over 10 times higher odds of not surviving compared to those without cancer.
   
5. BloodCarbonDioxide:

   - Odds Ratio: 0.085 (CI: 0.008-0.55)
   
   - Lower blood carbon dioxide levels significantly reduce the odds of not surviving.
  
6. BloodPH:

   - Odds Ratio: 8.07 (CI: 1.4-53.60)
   
   - Lower blood pH levels significantly increase the odds of not surviving.

### Conclusion:

The selected model indicates that factors such as concsiousness level, type of admission, age, cancer, blood carbon dioxide, and blood pH are significant predictors of survival. Patients who are unconscious, have acute admissions, are older, have cancer, and have abnormal blood gas levels are at higher risk of not surviving. These results can help identify high-risk patients and improve treatment strategies to increase survival rates.


## Question 2

How well does your chosen model fit the data? In assignment 3, deviance was used to assess model fit. However, for individual-level data, deviance is unsuitable. Instead, perform the Hosmer-Lemeshow goodness-of-fit test using the recommended R code.

### Approach:

- Understand the Hosmer-Lemeshow Test:

  - The Hosmer-Lemeshow test evaluates wether the observed event rates mateches the expected probabilities predicted by the model.
  
  - The null hypothesis is that the model fits the data well (a high p-value suggests no evidence of poor fit).

- Implementation

  - We have ot use te function for the test ResourceSelection::hoslem.test()
  
  - We have to specify the predicted probabilities from the final model and the actual outcomes. (m_step and Survival respectively).
  
  - Set the number of groups for grouping observation 
  
### Code and results :
```{r echo=FALSE}
# Calculate predicted probabilities
predicted_probs <- predict(m_step, type = "response")

# Perform Hosmer-Lemeshow goodness-of-fit test
hoslem_test <- ResourceSelection::hoslem.test(data_ca4$Survival, predicted_probs, g = 10)

# Display test results
hoslem_test
```

**Hosmer-Lemeshow Test Results:**

  - Test Statistic ($\chi^2$): 3.2956
  
  - Degrees of Freedom: 8
  
  - p-value: 0.9145
  
**Interpretation:**

  1. The p-value of 0.9145 tells us not reject the null hypothesis that the model fits the data well. 
  
### Conclusion:

The Hosmer-Lemeshow test indicates that the model fits the data well. The observed event rates are consistent with the expected probabilities predicted by the model. This suggests that the model is a good fit for the data and can be used to make accurate predictions about survival probabilities.


