---
title: "AssignmentIV"
author: "Silpa Soni Nallacheruvu (19980824-5287) Hernan Aldana (20000526-4999)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

# Exercise 4:1

##  Question 1

Identify factors that affect the probability of not surviving.

### Approach:

- Prepare the data by combining categories fo selected variables.

- Fit an empty model (intercept only) and a full model (all predictors).

- Perform a step wise variable selection to identify the best model.

- Identify significant predictors and compute the odds ratios and confidence intervals.

- Interpret the results.

### Code and results :

```{r echo=FALSE}
# Data Preparation
data4$v5[data4$v5 > 1] <- 0  # Combine categories for v5

# Empty and Full Models
m_empty <- glm(v2 ~ 1, family = binomial, data = data4)  # Empty model
m_full <- glm(v2 ~ ., family = binomial, data = data4)   # Full model

# Stepwise Selection (AIC)
mstep3_AIC <- step(m_full, direction = "both", trace = FALSE)  # Both directions

# Summary of Final Model
summary(mstep3_AIC)

# Odds Ratios and Confidence Intervals
odds_ratios <- exp(coef(mstep3_AIC))
conf_intervals <- exp(confint(mstep3_AIC))

# Create a table
library(knitr)
summary_table <- data.frame(
  Predictor = names(odds_ratios),
  Odds_Ratio = odds_ratios,
  CI_Lower = conf_intervals[, 1],
  CI_Upper = conf_intervals[, 2]
)
kable(summary_table, caption = "Odds Ratios and Confidence Intervals for Final Model")
```

### Conclusion :

** Model selection **

- Through stepwise variable selection using AIC, to identify the best model for predicting survival. The final model included the following variables as significant predictors for survival:

  - v1 (Patient):
  
  - v3 (Age):
  
  - v7 (Heart_lung_treatment)
  
  - v14








```{r echo=FALSE}
# Empty model (intercept only)
m_empty <- glm(v2 ~ 1, family = binomial, data = data4)

# Full model (all predictors)
m_full <- glm(v2 ~ ., family = binomial, data = data4)

# Display the summary of the full model to check all predictors
summary(m_full)
```

```{r echo=FALSE}
# Perform stepwise variable selection
library(MASS) # for stepAIC function
stepwise_model <- stepAIC(logistic_model, direction = "both")

# Summarize the final model
summary(stepwise_model)

```

## Output : 

## Observation : 


