---
title: "AssignmentIII"
subtitle: "Log-Linear Models"
author: "Silpa Soni Nallacheruvu (19980824-5287) Hernan Aldana (20000526-4999)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(knitr)
library(stringr)
```

---

# Exercise 3:1 (Higher Dimension Table)

## Question1: 

Fit several models in order to find a 'good' model for the given data collected from a birth clinic, which includes information on the mother's age, her smoking habits (number of cigarettes per day), gestational age (in days) and the survival status of the child.

```{r echo=FALSE, message=FALSE}
# Create the table data
data <- data.frame(
  `Mother's age` = c("< 30", "< 30", "< 30", "< 30", "30+", "30+", "30+", "30+"),
  `Smoking habits` = c("< 5", "< 5", "5+", "5+", "< 5", "< 5", "5+", "5+"),
  `Gestational age` = c("< 260", ">= 260", "< 260", ">= 260", "< 260", ">= 260", "< 260", ">= 260"),
  `Child survival - No` = c(50, 24, 9, 6, 41, 14, 4, 1),
  `Child survival - Yes` = c(315, 4012, 40, 459, 147, 1594, 11, 124)
)

# Generate the table
kable(data, caption = "Data from the study on the association of variables with child survival", col.names = c("Mother's age", "Smoking habits", "Gestational age", "Child survival - No", "Child survival - Yes"))
```

### Approach: 

**1. Read the data:**

- The given dataset 'data_ca3.csv' contains the variables X, Y, Z, and V, along with their corresponding frequencies (n).
  
**2. Fit a saturated model:**

- Fit a saturated model which includes all four variables X, Y, Z, and V and all their interactions. It fits the data perfectly and serves as the reference model.
  
**3. Reduced Models:**

- We started by removing the 4 way interaction term from the saturated model.
- Then, we removed the 3 way interaction terms, then 2 way interaction terms, and finally we fit a model with the only the main effects.
- We removed interactions in a systematic way, where higher order interactions are removed before lower order interactions to evaluate the effect of each interaction term on the model.
  
**4. Model Comparison:**

- We compared the models using deviance, degrees of freedom, p-value, and AIC.
- We calculated the p-value using the chi-square distribution calculated from the deviance and degrees of freedom.


### R Output:

```{r echo=FALSE, message=FALSE}

# Read the data
data3<-read.csv("data_ca3.csv")

# Create a data frame to store results
model_results <- data.frame(
  Model = character(),
  Deviance = numeric(),
  df = numeric(),
  `p-value` = numeric(),
  AIC = numeric()
)

# Define a function to extract model information
extract_model_info <- function(model, description) {
  data.frame(
    Model = description,
    Deviance = model$deviance,
    df = model$df.residual,
    `p-value` = 1 - pchisq(model$deviance, model$df.residual),
    AIC = AIC(model)
  )
}

# Saturated model:
msat<-glm(n~x*y*z*v, family=poisson(link=log), data=data3)

# Add the saturated model to results
model_results <- rbind(model_results, extract_model_info(msat, "XYZV"))

#generate interactions
terms <- c("x*y*z", "x*y*v", "x*z*v", "y*z*v")
interaction_combinations34 <- combn(terms, 4, simplify = FALSE)
interaction_combinations33 <- combn(terms, 3, simplify = FALSE)
interaction_combinations32 <- combn(terms, 2, simplify = FALSE)
interaction_combinations31 <- combn(terms, 1, simplify = FALSE)
terms2 <- c("x*y", "x*z", "x*v", "y*z", "y*v", "z*v")
interaction_combinations26 <- combn(terms2, 6, simplify = FALSE)
interaction_combinations25 <- combn(terms2, 5, simplify = FALSE)
interaction_combinations24 <- combn(terms2, 4, simplify = FALSE)
interaction_combinations23 <- combn(terms2, 3, simplify = FALSE)
interaction_combinations22 <- combn(terms2, 2, simplify = FALSE)
interaction_combinations21 <- combn(terms2, 1, simplify = FALSE)
terms3 <- c("x", "y", "z", "v")
interaction_combinations14 <- combn(terms3, 4, simplify = FALSE)
#combine all interactions
interaction_combinations <- c(interaction_combinations34, interaction_combinations33, interaction_combinations32, interaction_combinations31, interaction_combinations26, interaction_combinations25, interaction_combinations24, interaction_combinations23, interaction_combinations22, interaction_combinations21, interaction_combinations14)

# Function to generate and evaluate models with three-way interactions
evaluate_interactions <- function(data, terms, model_results) {
  # Loop through each combination of interactions
  for (combo in terms) {
    # Generate model formula
    interaction_terms <- paste(combo, collapse = "+")
    formula <- as.formula(paste("n ~", interaction_terms))
    
    # Fit the model
    model <- glm(formula, family = poisson(link = "log"), data = data)
    
    # Append results
    model_results <- rbind(
      model_results,
      extract_model_info(model, str_to_upper(paste(gsub("\\*", "", combo), collapse=',')))
    )
  }
  
  return(model_results)
}

model_results <- evaluate_interactions(data3, interaction_combinations, model_results)
#print the model results into a table
kable(model_results, caption = "Model Comparison Results", align = c("l", "c", "c", "c", "c"), col.names = c("Model", "Deviance", "df", "p-value", "AIC"))
```

### Conclusion:

**1. Trends in AIC Across Models:**

  - Models with only main effects or single two-way interactions have high AICs, indicating poor fit.
  - Adding specific two-way and three-way interactions significantly improves the AIC, highlighting the importance of these interaction terms for explaining the data.

**2. Impact of Higher-Order Interactions:**

  - Models including higher-order interactions (e.g., XYZV) exhibit extremely low p-values, suggesting overfitting and unnecessary complexity. The saturated model achieves perfect fit but at the cost of increased complexity, as reflected in its AIC.
  - Models with XYV, XZV as interactions have the highest p-values, indicating that removing some of the three-way interactions does not significantly reduce the model fit. These models are candidates for simpler yet statistically robust options.
  
  
## Question2:   

### Approach: 

To answer question 2, we aim to identify a good model that balances simplicity and fit. A good model should:

  - Explain the Relationships: Capture significant associations between variables.
  
  - Avoid Overfitting: Include only necessary interactions to prevent overfitting.
  
  - Optimize Fit: Minimize AIC and retain a good fit as assessed by likelihood ratio tests.

We assume that higher-order interactions (e.g., 3-way and 4-way) might not significantly contribute to explaining the relationships, as they may be too complex and difficult to interpret, as well as prone to overfitting. Therefore our focus will be on identifying the best model with two-way interactions. To do this we have to follow this steps:

  - Start with the Saturated Model: Fit a saturated model with all interactions to establish a baseline.
  
  - Fit Reduced Models: Gradually remove higher-order interactions (3-way and 4-way) and evaluate the impact on model fit.
  
  - Evaluate models: Compare models based on AIC, p-values, and goodness-of-fit, and p-values from LRT. 
  
  - Select the Best Model: Identify the model with the lowest AIC that balances fit and complexity.
  
```{r echo=FALSE, message=FALSE}
# Load the data
data3 <- read.csv("data_ca3.csv")

# Saturated model with all interactions
msat <- glm(n ~ x * y * z * v, family = poisson(link = "log"), data = data3)

# Reduced model with two-way interactions: XY, XZ, XV, YV, ZV
m1 <- glm(n ~ x * y + x * z + x * v + y * v + z * v, family = poisson(link = "log"), data = data3)

# Compare models using AIC and LRT
aic_results <- AIC(msat, m1)
lrt_results <- anova(msat, m1, test = "LRT")

# Output results
print("AIC Results:")
print(aic_results)

print("Likelihood Ratio Test Results:")
print(lrt_results)

# Summary of the chosen model (m1)
summary(m1)
```
### Results

**AIC Results:**

Saturate model (msat): AIC =  123.9732
Reduced model (m1): AIC =  113.7953

The reduced model (m1) has a lower AIC value compared to the saturated model, indicating a better balance between goodness-of-fit with fewer parameters.

**Residual Deviance:**

Saturation model (msat): 0.0000 (perfect fit)
Reduced model (m1): 1.8221

**Model coefficients**

All coefficients in the reduced model (m1) are statistically significant (p < 0.05), with the exception of a few borderline cases.

The chosen model(m1) includes significant two-way interactions XY, XZ, XV, YV, ZV. This interactions highlight the relationships between:

  - Mother's age and smoking habits (XY): Indicates that smoking habits vary significantly accross maternal age groups.
  
  - Mother's age and gestational age (XZ): Suggests that gestational age may be influenced by maternal age.
  
  - Mother's age and child survival (XV): Indicates that child survival is critically dependent on maternal age.
  
  - Smoking habits and child survival (YV): Shows the direct impact of smoking habits on child survival rates.
  
  - Gestational age and child survival (ZV): Reinforces that gestational age is a key factor in determining child survival rates.


### Conclusion:

The reduced model (m1) with interactions XY, XZ, XV, YV, ZV was selected as the best model because:

  - It has the lowest AIC.

  - It retains statistically significan two way interactions that provide interpretable results inot the relationships between variables.
  
  - Removing higher-order interactions (3-way and 4-way) did not significantly impact the model fit, as evidenced by the residual deviance and p-values.
  
  
  
  
  
### this fits better in question 2 rather as a conclusion in question 1

The model with the lowest AIC is the one with interactions XY, XZ, XV, YV, ZV, suggesting it balances goodness-of-fit and model complexity most effectively. This model includes all two-way interactions except YZ, capturing significant dependencies among variables while avoiding overfitting.  

Recommendation for a ‘Good’ Model:
Based on the results, the model with XY, XZ, XV, YV, ZV is recommended for its optimal balance of simplicity and fit (lowest AIC). However, models like XYV, XZV might also be considered if further parsimony is desired, given their statistical significance.
 

