---
title: "AssignmentI"
author: "Silpa Soni Nallacheruvu (19980824-5287) Hernan Aldana (20000526-4999)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(MASS)
library(epitools)
```

# Summary

This report analyzes opinions on legal abortion among men and women using data from a survey. The responses were summarized in a two-way table, and statistical methods, including percentages, hypothesis testing, odds ratios, and risk ratios, were applied to gain insights. The results are presented with detailed calculations and interpretations.

# Exercise 1:1

## Question 1: Percentage in Favor and Against Legal Abortion by Gender

### Approach : 

The percentages for being “in favor” and “against” legal abortion for men and women were calculated using the following formulas :

- Percentage (In Favor) :

${\text{Percentage (In Favor)} = \frac{\text{Count (In Favor) of the Gender}}{\text{Total Count of the Gender}} \times 100}$

- Percentage (Against) :

${\text{Percentage (Against)} = \frac{\text{Count (Against) of the Gender}}{\text{Total Count of the Gender}} \times 100}$

### Calculations : 
```{r percentage-calculations, echo=TRUE}
# Calculate percentages
women_in_favor <- 309
women_total <- 500
men_in_favor <- 319
men_total <- 600
women_against <- women_total - women_in_favor
men_against <- men_total - men_in_favor

# Percentages
women_in_favor_pct <- round((women_in_favor / women_total) * 100, 1)
women_against_pct <- round((women_against / women_total) * 100, 1)
men_in_favor_pct <- round((men_in_favor / men_total) * 100, 1)
men_against_pct <- round((men_against / men_total) * 100, 1)
```

### Summary : 

```{r percentage-summary, echo=FALSE}
# Create summary table
summary_table <- data.frame(
  Gender = c("Women", "Men"),
  `Percentage In Favor` = c(women_in_favor_pct, men_in_favor_pct),
  `Percentage Against` = c(women_against_pct, men_against_pct)
)

kable(summary_table, col.names = c("Gender", "% In Favor", "% Against"),
      caption = "Percentage In Favor and Against Legal Abortion by Gender", align = c("l", "c", "c"))
```


### Analysis : 

- Women: 61.8% are in favor, while 38.2% are against legal abortion.
- Men: 53.2% are in favor, while 46.8% are against legal abortion.

## Question 2: Hypothesis Testing (Pearson’s (X^2) and Likelihood Ratio (G^2))

### Approach : 

#### 1. Hypotheses : 

- Null Hypothesis (${H_0}$) : There is no difference in opinions between men and
women on legal abortion.

- Alternative Hypothesis (${H_A}$) : There is a difference in opinions between
men and women on legal abortion.

#### 2. Expected Counts : 

Under ${H_0}$, expected counts are calculated using: 

${E_{ij} = \frac{\text{i-th Row Total} \times \text{j-th Column Total}}{\text{Grand Total}}}$

#### 3. Pearson’s Chi-Squared Statistic :

The formula for Pearson’s Chi-Squared Statistic (${X^2}$) is: 

${X^2 = \sum_{i=1}^{r} \sum_{j=1}^{c} \frac{(O_{ij} - E_{ij})^2}{E_{ij}}}$

#### 4. Likelihood Ratio Statistic : 

The formula for Likelihood Ratio Statistic (${G^2}$) is: 

${G^2 = 2 \sum_{i=1}^{r} \sum_{j=1}^{c} O_{ij} \log\left(\frac{O_{ij}}{E_{ij}}\right)}$

where ${O_{ij}}$ is the Observed Count, ${E_{ij}}$ is the Expected Count 
of (i,j) cell with i-th index of X and j-th index of Y, r is the total number
of rows and c is the total number of columns.

### Calculations : 

```{r hypothesis-testing, echo=TRUE}
# Observed counts
observed <- matrix(c(309, 191, 319, 281), nrow = 2, byrow = TRUE)
rownames(observed) <- c("Women", "Men")
colnames(observed) <- c("In Favor", "Against")

# Totals
row_totals <- rowSums(observed)
col_totals <- colSums(observed)
grand_total <- sum(observed)

# Expected counts
expected <- outer(row_totals, col_totals) / grand_total

# Pearson's X^2
X2 <- sum((observed - expected)^2 / expected)

# Likelihood Ratio G^2
G2 <- 2 * sum(observed * log(observed / expected))

# Degrees of freedom
df <- (nrow(observed) - 1) * (ncol(observed) - 1)

# P-values
p_value_X2 <- pchisq(X2, df = df, lower.tail = FALSE)
p_value_G2 <- pchisq(G2, df = df, lower.tail = FALSE)


```

### Output : 

```{r echo=FALSE}
# Results
results <- data.frame(
  Statistic = c("Pearson's X^2", "Likelihood Ratio G^2"),
  Value = c(X2, G2),
  `P-value` = c(p_value_X2, p_value_G2)
)
kable(results, caption = "Hypothesis Testing Results", align = c("l", "c", "c"))
```

### Conclusion : 

We use a significance level ${\alpha}$ = 0.05 to test our null hypothesis ${H_0}$.
Both the ${X^2}$ and ${G^2}$ statistics yield small p-values, much lesser than 0.05. 
Hence, we reject the null hypothesis ${H_0}$, which indicates that there is a significant difference in opinions on legal abortion 
between men and women.

## Question 3: Odds Ratio and Confidence Interval

## Approach :

### The Odds Ratio :

The odds ratio (OR) is a measure used to quantify the strength of association 
between two categorical variables. 

#### Odds for Women:

${\text{Odds (Women)} = \frac{\text{In Favor (Women)}}{\text{Against (Women)}}}$

#### Odds for Men:

${\text{Odds (Men)} = \frac{\text{In Favor (Men)}}{\text{Against (Men)}}}$

#### Odds Ratio (OR) :

The odds ratio compares the odds of being “in favor” for men to that for women:

${\text{OR} = \frac{\text{Odds (Men)}}{\text{Odds (Women)}}}$

### Calculations : 

```{r odds-ratio, echo=TRUE}
# Odds calculation
odds_women <- women_in_favor / women_against
odds_men <- men_in_favor / men_against
# Odds ratio
odds_ratio <- odds_men / odds_women
```

### Output : 

```{r echo=FALSE}
# Results
results <- data.frame(
  Odds = c("Odds Women", "Odds Men", "Odds Ratio"),
  Value = c(odds_women, odds_men, odds_ratio)
)
kable(results, caption = "Odds Ratio Results", align = c("l", "c", "c"))
```

Alternative Odds Ratios can be formulated depending on the conditioning variable:

Odds ratio of being a man vs. a woman among those “in favor”:

${\text{OR} = \frac{\text{Men In Favor / Women In Favor}}{\text{Men Against / Women Against}}}$

Odds ratio of being against legal abortion for men vs. women:

${\text{OR} = \frac{\text{Men Against / Women Against}}{\text{Men In Favor / Women In Favor}}}$

Each odds ratio provides different insights based on the event being studied and 
the conditioning variable.

### Confidence Interval : 

To compute the 95% confidence interval for the odds ratio, let us use the 
logarithm of odds ratio for a more accurate result and then apply the delta method 
to get the confidence interval of odds ratio from the confidence interval 
of logarithm of odds ratio.

### Calculations : 

```{r echo=TRUE}
# Log odds ratio and standard error
log_odds_ratio <- log(odds_ratio)
se_log_odds <- sqrt(sum(1 / observed))

# Confidence interval
z <- qnorm(0.975)
ci_lower <- exp(log_odds_ratio - z * se_log_odds)
ci_upper <- exp(log_odds_ratio + z * se_log_odds)
```

### Output : 

```{r echo=FALSE}
# Results
odds_results <- data.frame(
  Measure = c("Odds Ratio", "95% CI Lower", "95% CI Upper"),
  Value = c(odds_ratio, ci_lower, ci_upper)
)
kable(odds_results, caption = "Odds Ratio and Confidence Interval", align = c("l", "c"))
```

###  Interpretation

The estimated odds ratio is 0.7015. This means that the odds of men being 
“in favor” of legal abortion are approximately 70% of the odds for women.
The 95% confidence interval (0.556, 0.886) does not include 1, indicating that 
the difference in odds is statistically significant at the 5% significance level. 
The estimated odds ratio indicates that the women are significantly more likely
to support legal abortion compared to men and the 95% confidence interval of 
the odds ratio support that indication. 

## Question 4: Risk Ratio and Confidence Interval

### Approach :

#### The Risk Ratio :

The risk ratio (also called relative risk, RR) quantifies the likelihood of an 
event occurring in one group relative to another group.

The risk refers to the probability of being “in favor” of legal abortion, 
calculated as:

${\text{Risk} = \frac{\text{Number of Individuals “In Favor”}}{\text{Total Number of Individuals in the Gender}}}$

#### Risk for Women:

${\text{Risk (Women)} = \frac{\text{Number of Women In Favor}}{\text{Total (Women)}}}$

#### Risk for Men:

${\text{Risk (Men)} = \frac{\text{Number of Men In Favor}}{\text{Total (Men)}}}$


#### Risk Ratio (Relative Risk) :

The risk ratio (RR) compares the risk of being “in favor” for men to that for women:

${\text{RR} = \frac{\text{Risk (Men)}}{\text{Risk (Women)}}}$

### Calculations : 

```{r risk-ratio, echo=TRUE}
# Risk calculation
risk_women <- women_in_favor / women_total
risk_men <- men_in_favor / men_total

# Risk ratio
risk_ratio <- risk_men / risk_women
```

### Output : 

```{r echo=FALSE}
results <- data.frame(
  Risk = c("Risk Women", "Risk Men", "Risk Ratio"),
  Value = c(risk_women, risk_men, risk_ratio)
)
kable(results, caption = "Risk Ratio Results", align = c("l", "c", "c"))
```

### Analysis : 

This value indicates that men are about 86.1% as likely as women to be 
“in favor” of legal abortion.

#### Confidence Interval :

To compute the 95% confidence interval for the risk ratio, let us use the 
logarithm of risk ratio for a more accurate result, similar to odds ratio and 
then apply the delta method to get the confidence interval of risk ratio from 
the confidence interval of logarithm of risk ratio.

The formula for standard error is:

${\text{SE} = \sqrt{\frac{1}{\text{In Favor (Men)}} - \frac{1}{\text{Total (Men)}} + \frac{1}{\text{In Favor (Women)}} - \frac{1}{\text{Total (Women)}}}}$

### Calculations : 

```{r echo=TRUE}
# Log risk ratio and standard error
log_risk_ratio <- log(risk_ratio)
se_log_risk <- sqrt(1 / women_in_favor - 1 / women_total + 1 / men_in_favor - 1 / men_total)

# Confidence interval
ci_lower_rr <- exp(log_risk_ratio - z * se_log_risk)
ci_upper_rr <- exp(log_risk_ratio + z * se_log_risk)

```

### Output : 

```{r echo=FALSE}
# Results
risk_results <- data.frame(
  Measure = c("Risk Ratio", "95% CI Lower", "95% CI Upper"),
  Value = c(risk_ratio, ci_lower_rr, ci_upper_rr)
)
kable(risk_results, caption = "Risk Ratio and Confidence Interval", align = c("l", "c"))
```

### Interpretation

The estimated risk ratio is 0.861. This means that men are about 86.1% as 
likely as women to be “in favor” of legal abortion. The 95% confidence interval 
(0.756, 0.980) does not include 1, indicating that the difference in risk is 
statistically significant at the 5% level. The estimated risk ratio indicates
that the women are significantly more likely to support legal abortion compared 
to men and the 95% confidence interval of the risk ratio support that indication.

### Comparison with Odds Ratio

- Odds Ratio:  Previously calculated as 0.7015, quantifying the odds of being 
            “in favor” for men compared to women.
            
- Risk Ratio: Calculated as 0.861, quantifying the relative probability of being “in favor.”

The odds ratio measures the ratio of odds, which can overstate the association, 
especially when the event probability is high (e.g., large proportions of people “in favor”).
The risk ratio provides a more intuitive interpretation as it measures the relative likelihood.


## Question 5: Verification of Previous Calculations

### Approach : 

#### 1. Generate Frequency Table and Calculate Row Percentages

Verify the calculations for percentages, hypothesis testing, odds ratio, and risk ratio from Questions 1–4 using built-in R functions. The results are compared with the manual calculations presented earlier.

```{r echo=FALSE}
#prepare the data and calculate row percentages
tab1<- as.table(rbind(c(309, 191), c(319, 281)))
dimnames(tab1) <- list(gender = c("women", "men"),opinion = c("favor","against"))
addmargins(tab1)
addmargins(prop.table(tab1,1),2)
```

#### 2. Calculate X2, G2 and p-values

```{r echo=FALSE}
# Calculate X2, G2 and p-values
chisq.test(tab1,correct=FALSE)
# install.packages("MASS")

log_likelihood_ratio <- loglm(~gender+opinion,tab1)
print(log_likelihood_ratio)
# The function loglm fits a loglinear model, but we leave the details about that.
# Log linear models will be handled in assignment 3 but with the function glm instead of loglm.
```
##### Comparison : 
- (X^2) statistic, degrees of freedom, and p-value displayed, matching the manual calculations.
- (G^2) statistic and p-value are also displayed, confirming the earlier results.

#### 3. Calculate Odds Ratio and Confidence Interval

```{r echo=FALSE}
# Calculate odds ratio and 95% confidence interval
# install.packages("epitools")
oddsratio(tab1, method = "wald", rev="neither")
```

reverse the rows : odds of women to men :
 
```{r echo=FALSE}
# To obtain the intended odds/risk ratio, you may need to reverse the rows or columns:
oddsratio(tab1, method = "wald", rev="row")
```

 reverse the column : odds of men to women :
 
```{r echo=FALSE}
# To obtain the intended odds/risk ratio, you may need to reverse the rows or columns:
oddsratio(tab1, method = "wald", rev="col")
```
reverse both : odds of women to men :

```{r echo=FALSE}
# To obtain the intended odds/risk ratio, you may need to reverse the rows or columns:
oddsratio(tab1, method = "wald", rev="both")
```

#### Comparison :

The odds ratio and its confidence interval are verified, and the results align with the manually calculated values when the appropriate rev option is used.


#### 4. Calculate Risk Ratio and Confidence Interval

```{r echo=FALSE}
# Calculate risk ratio with default settings
risk_ratio_default <- riskratio(tab1, rev = "neither")
print(risk_ratio_default)

# Check alternative row/column reversals to match the intended risk ratio
risk_ratio_row <- riskratio(tab1, rev = "row")
risk_ratio_col <- riskratio(tab1, rev = "col")
risk_ratio_both <- riskratio(tab1, rev = "both")
print(risk_ratio_row)
print(risk_ratio_col)
print(risk_ratio_both)
```

#### Comparison :

The risk ratio and its confidence interval are verified, and the results align with the manually calculated values when the appropriate rev option is used.


# Exercise 1:2

## Question 1:

### Approach : 

We are analyzing the admissions data from the University of California, Berkeley, following the 2x2 contingency table:

```{r echo=FALSE}

tab2 <- matrix(
  c(1198, 1493, 557, 1278), # Data in row-major order
  nrow = 2,                 # Two rows: Men and Women
  byrow = TRUE,             # Fill matrix by rows
  dimnames = list(
    gender = c("Men", "Women"),        # Row names
    admission = c("Admitted", "Not Admitted") # Column names
  )
)

# Display the table
knitr::kable(
  tab2,
  caption = "Summary Table"
)


```
Our goal is to perform the following analyses:

  1. Calculate the percentages of admitted and not admitted applicants separately.  
  
  2. Test for independence between gender and admission using:
  
      - Pearson’s Chi-Squared Test  
      - Likelihood Ratio Test
      
  3. Calculate the odds ratio and its 95% confidence interval.
  
  4. Calculate the risk ratio and its 95% confidence interval.  

1. Calculating Percentages

For men:

  - Total men: $n_{\text{Men}} = 2691$  
  
  - Admitted men: $a_{\text{Men}} = 1198$  
  
  - Not admitted men: $n_{\text{Men}} - a_{\text{Men}} = 1493$  

Percentages

$\text{Percentage Admitted (Men)} = \left( \frac{1198}{2691} \right) \times 100\% \approx 44.53\%$  

$\text{Percentage Not Admitted (Men)} = 100\% - 44.53\% = 55.47\%$

For women:  

  - Total women: $n_{\text{Women}} = 1835$  

  - Admitted women: $a_{\text{Women}} = 557$  

  - Not admitted women: $n_{\text{Women}} - a_{\text{Women}} = 1278$  
  
Percentages

$\text{Percentage Admitted (Women)} = \left( \frac{557}{1835} \right) \times 100\% \approx 30.35\%$  

$\text{Percentage Not Admitted (Women)} = 100\% - 30.35\% = 69.65\%$  

```{r echo=FALSE}
# Admissions data (Berkeley)
tab3 <- data.frame(
  Gender = c("Men", "Women"),
  `% Admitted` = c(44.53, 30.35),
  `% Not Admitted` = c(55.47, 69.65)
)
# Display the table
knitr::kable(
  tab3,
  col.names = c("Gender", "% Admitted", "% Not Admitted"),
  caption = "Summary Table",
  align = c("l", "c", "c")
)
```
2. Testing for Independence

Null hypothesis ($H_0$): Gender and admission are independent.  

Alternative hypothesis ($H_1$): There is an association between gender and admission Status.

Expected counts under $H_0$ are calculated as:  

$E_{ij} = \frac{(\text{Row Total}_i) \times (\text{Column Total}_j)}{\text{Grand Total}}$

  1. Men Admitted:  

      $E_{11} = \frac{R_1 \times C_1}{N} = \frac{2691 \times 1755}{4526} \approx 1043.46$  

  2. Men not Admitted:  

      $E_{12} = \frac{R_1 \times C_2}{N} = \frac{2691 \times 2771}{4526} \approx 1647.54$  

  3. Women Admitted:  

      $E_{21} = \frac{R_2 \times C_1}{N} = \frac{1835 \times 1755}{4526} \approx 711.54$. 

  4. Women not Admitted:  

      $E_{22} = \frac{R_2 \times C_2}{N} = \frac{1835 \times 2771}{4526} \approx 1123.46$. 

Calculating the Pearson's Chi-Squared Test Statistic:  

$\chi^2 = \sum_{i=1}^{2} \sum_{j=1}^{2} \frac{(O_{ij} - E_{ij})^2}{E_{ij}}$  

Calculating the Likelihood Ratio Test Statistic:  

$G^2 = 2 \sum_{i=1}^{2} \sum_{j=1}^{2} O_{ij} \ln\left( \frac{O_{ij}}{E_{ij}} \right)$  

where:  

  - $O_{ij}$: Observed counts  

  - $E_{ij}$: Expected counts  

The degrees of freedom are calculated for both statistics as $df=(r-1)(c-1)$  

3. Estimating the Odds Ratio

Calculating Odds  

  - Odds of Admission for Men:  

    $\text{Odds}_{\text{Men}} = \frac{\text{Admitted Men}}{\text{Not Admitted Men}}$  

  - Odds of Admission for Women:  

    $\text{OR} = \frac{\text{Odds}{\text{Men}}}{\text{Odds}{\text{Women}}}$  

Calculating the Standard Error of $\log(\text{OR})$:

$SE = \sqrt{ \frac{1}{a_{\text{Men}}} + \frac{1}{b_{\text{Men}}} + \frac{1}{a_{\text{Women}}} + \frac{1}{b_{\text{Women}}} }$

The 95% confidence interval for $\log(\text{OR})$ is calculated as:

$\ln(\text{OR}) \pm Z_{0.975} \times SE$

However we still have to get the confidence interval for the odds ratio by exponentiating the bounds.

4. Estimating the Risk Ratio

Calculating Risks (Probabilities)  

  - Risk of Admission for Men:  

    $P_{\text{Men}} = \frac{\text{Admitted Men}}{\text{Total Men}}$  

  - Risk of Admission for Women:  

    $P_{\text{Women}} = \frac{\text{Admitted Women}}{\text{Total Women}}$  

Calculating Risk Ratio (RR)  

$\text{RR} = \frac{P_{\text{Men}}}{P_{\text{Women}}}$  

Calculating the Standard Error of $\log(\text{RR})$:

$SE = \sqrt{ \frac{1 - P_{\text{Men}}}{a_{\text{Men}}} + \frac{1 - P_{\text{Women}}}{a_{\text{Women}}} }$

The 95% confidence interval for $\log(\text{RR})$ is calculated as:

${\ln(\text{RR}) \pm Z_{0.975} \times SE}$

However we still have to get the confidence interval for the risk ratio by exponentiating the bounds.

### Code

```{r echo=FALSE}
# Perform chi-squared test
chisq_result <- chisq.test(tab2, correct = FALSE)
chisq_result
```

```{r echo=FALSE}
tab2_table <- as.table(tab2)
# Calculate G2 using loglinear model
loglm_result <- loglm(~ gender + admission, tab2_table)
loglm_result
```
```{r echo=FALSE}
# Calculate Odds Ratio and Confidence Interval
odds_ratio_result <- oddsratio(tab2, method = "wald", rev = "neither")
odds_ratio_result
```

```{r echo=FALSE}
# Calculate Risk Ratio and Confidence Interval
risk_ratio_result <- riskratio(tab2, rev = "neither")
risk_ratio_result
```

### Results and Interpretation

1. Chi-Squared Test:

${\chi^2 = 92.205, p < 2.2e^{-16}}$

The large chi-squared statistic and the small p-value indicate a significant association between gender and admission status. Therefore rejecting the null  hypothesis of independence.

2. Likelihood Ratio Test:

${G^2 = 93.98, p < 0.001}$

Similarly as the chi-squared test, the likelihood ratio test also indicates a significant association between gender and admission.

3. Odds Ratio:

${\text{OR} = 1.841, 95\% \text{ CI } (1.624, 2.087)}$

According to the odds ratio, men have 1.841 times higher odds of being admitted compared to women. The 95% confidence interval does not include 1, indicating a statistically significant difference in admission odds.

4. Risk Ratio:

${\text{RR} = 1.467, 95\% \text{ CI } (1.353, 1.591)}$

Men are 1.255 times more likely to be admitted compared to women. The confidence interval does not include 1, indicating a statistically significant difference in admission risk.

### Conclusion

There is a statistically significant association between gender and admission status at the University of California, Berkeley during the year of 1975. Men have higher odd and higher probability of admission compared to women. 

## Question 2:

### Approach :

The approach is similar to the previous question. We will analyze what happens if we replace all numbers in the table with one tenth of their original values and perform the same analysis, which we will compare with the previous results.

### Code:

```{r echo=FALSE}
# Function to analyze data
analyze_table <- function(data) {
  # Perform Chi-Squared Test
  chi_result <- chisq.test(data, correct = FALSE)
  
  # Odds Ratio and Risk Ratio
  odds_ratio_result <- oddsratio(data, method = "wald", rev = "neither")
  risk_ratio_result <- riskratio(data, rev = "neither")
  
  list(
    "Chi-Squared Test" = chi_result,
    "Likelihood Ratio" = loglm(~gender + admission, data),
    "Odds Ratio" = odds_ratio_result$measure,
    "Risk Ratio" = risk_ratio_result$measure
  )
}

# Original analysis
original_result <- analyze_table(tab2)

# Replace all numbers in the table by one-tenth
tab2_tenth <- round(tab2 / 10, 0)
tenth_result <- analyze_table(tab2_tenth)
one_hundreth_result <- analyze_table(round(tab2 / 100, 0))

# Print results for comparison
list(
  "Original" = original_result,
  "One-Tenth" = tenth_result,
  "One-Hundreth" = one_hundreth_result
)
```

### Results and Interpretation:

1. Chi-Squared Test and Likelihood Ratio Test:

The Chi-Squared  $\chi^2$ and Likelihood Ratio $G^2$ statistics, as well as their associated p-values, measure the association between gender and admission. The results for the original, one-tenth, and one-hundredth tables are summarized below:

```{r echo=FALSE}
# Create the data frame for the table
result_table <- data.frame(
  Dataset = c("Original", "One-Tenth", "One-Hundredth"),
  Chi2_Statistic = c(92.205, 9.241, 0.775),
  G2_Statistic = c(93.449, 9.364, 0.783),
  p_value = c("< 2.2 × 10^-16", "0.002", "0.379")
)

# Print the table using knitr::kable
knitr::kable(
  result_table,
  col.names = c("Dataset", "χ² Statistic", "G² Statistic", "p-value"),
  caption = "Summary of Chi-Squared and Likelihood Ratio Tests",
  align = c("l", "c", "c", "c")
)
```

As it can be seen, both the Chi-Squared and Likelihood Ratio tests show a dramatic decrease as the table is scaled down, reflecting thus the reduced sample size.
On the other hand, the p-values show a completely different behavior. While the p-value for the original table is extremely small, indicating a significant association between gender and admission, as the table is scaled down, the p-value increases, suggesting that the association cannot be reliably detected with such a small sample size.

2. Odds Ratio:

The odds ratio quantifies the strength of the association between gender and admission. The results for the original, one-tenth, and one-hundredth tables are summarized below:

```{r echo=FALSE}
# Create the data frame for the Odds Ratio table
odds_ratio_table <- data.frame(
  Dataset = c("Original", "One-Tenth", "One-Hundredth"),
  `Odds Ratio (Women)` = c(1.841, 1.841, 1.733),
  `95% Confidence Interval` = c("[1.624, 2.087]", "[1.240, 2.734]", "[0.507, 5.928]")
)

# Print the Odds Ratio table using knitr::kable
knitr::kable(
  odds_ratio_table,
  col.names = c("Dataset", "Odds Ratio (Women)", "95% Confidence Interval"),
  caption = "Odds Ratio for Women Across Datasets",
  align = c("l", "c", "c")
)
```

The odds ratio across the three datasets shows a consistent value, varied only slightly. The confidence interval, however, widens as the sample size decreases, reflecting the increased uncertainty in the estimate. So for the original dataset, the odds ratio is 1.841, with a 95% confidence interval of [1.624, 2.087], which is very narrow, reflecting high precision. On the other hand, for the one-hundredth dataset, the odds ratio is 1.733, with a 95% confidence interval of [0.507, 5.928], which is much wider, reflecting low precision.

3. Risk Ratio:

The risk ratio compares the probability of admission for men and women. The results for the original, one-tenth, and one-hundredth tables are summarized below:

```{r echo=FALSE}
# Create the data frame for the Risk Ratio table
risk_ratio_table <- data.frame(
  Dataset = c("Original", "One-Tenth", "One-Hundredth"),
  `Risk Ratio (Women)` = c(1.255, 1.256, 1.232),
  `95% Confidence Interval` = c("[1.200, 1.314]", "[1.088, 1.450]", "[0.781, 1.941]")
)

# Print the Risk Ratio table using knitr::kable
knitr::kable(
  risk_ratio_table,
  col.names = c("Dataset", "Risk Ratio (Women)", "95% Confidence Interval"),
  caption = "Risk Ratio for Women Across Datasets",
  align = c("l", "c", "c")
)
```


The risk ratio across the three datasets shows a consistent value, varied only slightly. The confidence interval, however, widens as the sample size decreases, reflecting the increased uncertainty in the estimate. So for the original dataset, the risk ratio is 1.255, with a 95% confidence interval of [1.200, 1.314], which is very narrow, reflecting high precision. On the other hand, for the one-hundredth dataset, the risk ratio is 1.232, with a 95% confidence interval of [0.781, 1.941], which is much wider, reflecting low precision. Moreover, for the original dataset, the confidence interval does not include 1, indicating statistical significance. However, for the one-hundredth dataset, the confidence interval includes 1, indicating no statistical significance.

### Conclusion

This question helps us highlight the importance of a sufficiently large sample size for reliabale statistical inference. While relative measures such as odds and risk ratios remain relatively stable across different sample sizes, the precision of the estimates decreases as the sample size decreases. In addition, absolute measures such as $\chi^2$ and $G^2$ lose significance as the sample size decreases.

## Question 3: